---
title: "Developing Data Products Week 3 Assignment"
author: "Joshua Huber"
date: "June 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
set.seed(12345)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

## Problem Statement
We would like to display an interactive [choropleth map](https://en.wikipedia.org/wiki/Choropleth_map) showing the states of the US and how many hate crimes are reported in each state during the most recent year reported (currently 2016).

## Data acquisition
We start by getting two data sets from a federal government website. The data comes in the form of an Excel file, so we use the `xlsx` package to read the file.

```{r, message=FALSE}
if(!file.exists("Agency Hate Crime Reporting by State, 2016.xls")){
    download.file(  "https://ucr.fbi.gov/hate-crime/2016/tables/table-12/table_12_agency_hate_crime_reporting_by_state_2016.xls/output.xls"
                  , "Agency Hate Crime Reporting by State, 2016.xls")
}
library(xlsx)
hate_crimes <- read.xlsx('Agency Hate Crime Reporting by State, 2016.xls', 1, stringsAsFactors=FALSE)
```

## Data Cleaning and Processing
We subset the data, keeping only the columns that contain the state name and the number of incidents. We also keep only the rows that contain the per-state data, removing rows that contain summary and unnecessary information.  Since we are going to make a plotly geo map, we need to add the state abbreviation (code).
```{r}
hate_crimes <- hate_crimes[4:53, c(1,5)]
names(hate_crimes) <- c('State', 'Incidents_Reported')
hate_crimes$Incidents_Reported <- as.numeric(hate_crimes$Incidents_Reported)
hate_crimes$code <- c('AL', 'AK', 'AZ', 'AR', 'CA', 'CO', 'CT', 'DE', 'FL', 'GA', 'HI', 'ID', 'IL', 'IN', 'IA', 'KS', 'KY', 'LA', 'ME', 'MD', 'MA', 'MI', 'MN', 'MS', 'MO', 'MT', 'NE', 'NV', 'NH', 'NJ', 'NM', 'NY', 'NC', 'ND', 'OH', 'OK', 'OR', 'PA', 'RI', 'SC', 'SD', 'TN', 'TX', 'UT', 'VT', 'VA', 'WA', 'WV', 'WI', 'WY')
```

## Interactive Map

```{r, message=FALSE, out.width='100%'}
library(plotly)

# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)

# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

plot_geo(hate_crimes, locationmode = 'USA-states') %>%
  add_trace(
    z = ~Incidents_Reported, text = ~State, locations = ~code,
    color = ~Incidents_Reported, colors = 'Reds'
  ) %>%
  colorbar(title = "Incidents") %>%
  layout(
    title = '2016 Hate Crimes Reported by State',
    geo = g
  )
```
